---
title: "Where do Bike/Ped Crashes Occur?"
author: "Kyu Ri Kim"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
library(car)  #for scatter plot matrix
library(MASS) # glm and glm.nb
library(lattice) #for the levelplots
library(PerformanceAnalytics)
library(grid)
library(gridExtra)
library(readr) #for read data
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(knitr)
library(ggplot2)
library(ggcorrplot)
library(plotly)
library(dplyr)
library(lubridate)
library(leaflet)
library(sf)
library(htmltools)


# data (CVS) included
data_segment_ave <- read_csv("G:/My Drive/8_TREC/bikeped/data/data_6yrcrash_vol.csv")
data_state_year <- read_csv("G:/My Drive/8_TREC/bikeped/data/data_volume_crash_year_new.csv")
data_state_month <- read_csv("G:/My Drive/8_TREC/bikeped/data/data_volume_crash_month.csv")
speed <- read_csv("G:/My Drive/8_TREC/bikeped/data/speed_id_halfmi_month.csv")

bikecrash_segment_year <- read_csv("G:/My Drive/8_TREC/bikeped/data/bikecrash_long.csv")
pedcrash_segment_year <-read_csv("G:/My Drive/8_TREC/bikeped/data/pedcrash_long.csv")
p_monthly_bike <- read_csv("G:/My Drive/8_TREC/bikeped/data/p_monthly_bike.csv")
p_monthly_ped <- read_csv("G:/My Drive/8_TREC/bikeped/data/p_monthly_ped.csv")


# data (CVS) included
sf_bikevol <- read_sf("G:/My Drive/8_TREC/bikeped/data/bikevol.shp") %>% mutate(segment_area_id=segment_ar) 
sf_bikevol <- left_join(sf_bikevol,data_segment_ave,by='segment_area_id')
sf_bikevol_half_or <- read_sf("G:/My Drive/8_TREC/bikeped/data/or_bikevol_half.shp")

sf_pedvol <- read_sf("G:/My Drive/8_TREC/bikeped/data/bikevol.shp")%>% mutate(segment_area_id=segment_ar) 
sf_pedvol <- left_join(sf_pedvol,data_segment_ave,by='segment_area_id')


sf_bikepedvol_vol_dcva <- read_sf("G:/My Drive/8_TREC/bikeped/data/bikepedvol_half_dcva.shp")

sf_bikecrash_dcva <- read_sf("G:/My Drive/8_TREC/bikeped/data/bikecrash_dcva.shp")
sf_bikecrash_dcva_volume <- read_sf("G:/My Drive/8_TREC/bikeped/data/bike_crash_dcva_forvolumecase.shp")
sf_pedcrash_dcva <- read_sf("G:/My Drive/8_TREC/bikeped/data/pedcrash_dcva.shp")
sf_pedcrash_dcva_volume <- read_sf("G:/My Drive/8_TREC/bikeped/data/ped_crash_dcva_forvolumecase.shp")
sf_popdenhalf_dcva <- read_sf("G:/My Drive/8_TREC/bikeped/data/dcvapopden_half.shp")
sf_bikecrash_or <- read_sf("G:/My Drive/8_TREC/bikeped/data/bikecrash_or.shp")
sf_bikecrash_or_volume <- read_sf("G:/My Drive/8_TREC/bikeped/data/bike_crash_or_forvolumecase.shp")
sf_popdenhalf_or <- read_sf("G:/My Drive/8_TREC/bikeped/data/orpopden_half.shp")


# create pivot table for monthly bike volume
bikevol_segment_month <- mutate(p_monthly_bike, id_yr=paste(segment_area_id,'_',year))
bikevol_segment_month <- mutate(p_monthly_bike, id_yrmon=paste(segment_area_id,'_',year,'_',month))


# create pivot table for monthly ped volume                                
pedvol_segment_month <- mutate(p_monthly_ped, id_yr=paste(segment_area_id,'_',year))
pedvol_segment_month <- mutate(p_monthly_ped, id_yrmon=paste(segment_area_id,'_',year,'_',month))


# create pivot table for monthly speed by segment_id by year & month    
speed_state_month <- group_by(speed, state, month)
speed_state_month <- summarise(speed_state_month,
                        speed_50th=mean(segment_50thmed),
                        speed_85th=mean(segment_85thmed),
                        speed_max=max(segment_max))


# create pivot table for yearly bike volume
data_bike_year <- group_by(bikevol_segment_month, segment_area_id, state, year)
bikevol_segment_year <- summarise(data_bike_year,
                                  bikevol_total=sum(vol_mdaily),
                                  n=n(),
                                  bikevol=sum(vol_mdaily)/n)

bikevol_segment_year <- mutate(bikevol_segment_year, id_yr=paste(segment_area_id,'_',year))

# create pivot table for yearly ped volume
data_ped_year <- group_by(pedvol_segment_month, segment_area_id, state, year)
pedvol_segment_year <- summarise(data_ped_year,
                                  pedvol_total=sum(vol_mdaily),
                                  n=n(),
                                  pedvol=sum(vol_mdaily)/n)

pedvol_segment_year <- mutate(pedvol_segment_year, id_yr=paste(segment_area_id,'_',year))


bikecrash_segment_year <- mutate(bikecrash_segment_year, id_yr=paste(segment_area_id,'_',year))
pedcrash_segment_year <- mutate(pedcrash_segment_year, id_yr=paste(segment_area_id,'_',year))

# create pivot table for yearly speed by segment_id by year & month    
speed_segment <- group_by(speed, segment_area_id, year)
speed_segment <- summarise(speed_segment,
                        speed_50th=mean(segment_50thmed),
                        speed_85th=mean(segment_85thmed),
                        speed_max=max(segment_max))
speed_segment <- mutate(speed_segment, id_yr=paste(segment_area_id,'_',year))


# merge bike/ped volume data and speed data by id(segment_area_id, year, month)
bikevol_speed_segment_year <- inner_join(bikevol_segment_year,speed_segment,by='id_yr')
bikevol_crash_speed_segment_year <- inner_join(bikevol_speed_segment_year,bikecrash_segment_year,by='id_yr')

pedvol_speed_segment_year <- inner_join(pedvol_segment_year,speed_segment,by='id_yr')
pedvol_crash_speed_segment_year <- inner_join(pedvol_speed_segment_year,pedcrash_segment_year,by='id_yr')


```

# 1. Introduction

## 1.1. Purpose
- **Exploring the Relationship Between Volumes, Crashes, and Speeds**

## 1.2. Unit for Analysis
- **Space within a half-mile straight-line buffer from the permanent count location**

- **Census Block Group (for Population Density)**

# 2. Data

## 2.1. Basic Information of Data

- **Location**: Oregon, Virginia, Washington D.C.

- **Analysis Year**: 2018-2023

- **Used Data Type**: Bike & Pedestrian count (volume), Crash , Census Block group, Land Use, Speed

- **Data Source/Year**

|               Data                |                 Source                |    Year    |
|:---------------------------------:|:-------------------------------------:|:----------:|
| Bike/Pedestrian Counts            |       BikePed Portal                  | 2018--2023 |
| Bike/Pedestrian--involved Crash   |       ODOT*, VDOT, DDOT               | 2018--2023 |
| Land Use (Zoning) Type--DC        |       Open Data DC                    |    2016    |
| Land Use (Zoning) Type--OR        |       Oregon GEOHub Data              |    2023    |
| Land Use (Zoning) Type--VA        |       Arlington Virginia ACMaps       |    2021    |
| Census Block Group Population & Land Area--DC | Open Data DC              |    2020    |
| Census Block Group Population & Land Area--OR | Oregon GEOHub Data        |    2020    |
| Census Block Group Population & Land Area--VA | Virginia Open Data Portal |    2020    |
| Speed--DC, OR, VA                 |       RITIS                           | 2018--2023 |

*Note that ODOT provides crash data with location information through 2022. Since location information is a very important feature in this study, only cases where the location of each crash case was provided were analyzed.

## 2.2. Daily Bike/Ped Volume Data Count by State: 2018-2023

Following characteristics of pedestrian and bike volume and crash counts are accumulated of each permanent count location

```{r echo=FALSE, warning=FALSE}

p1<- ggplot(data_segment_ave) +
  aes(x=state, y=vol_6ydaily_bike, color = state) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
  theme_light()+
  theme(
    axis.title.y=element_blank(),
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Daily Bike Counts") +
  
  xlab("State")
  

ped_vol_state <- data_segment_ave %>% filter(state !="OR")
p2<-ggplot(ped_vol_state) +
  aes(x=state, y=vol_6ydaily_ped, color = state) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
 theme_light()+
   theme(
    axis.title.y=element_blank(),
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Pedestrian Counts") +
    xlab("State")

grid.arrange(p1, p2, nrow = 1)
```

```{r echo=FALSE, warning=FALSE}
# Prepare the text for the tooltip:
pop_bike <- paste(
   "State: ", sf_bikevol$state, "<br/>", 
   "Segment: ", sf_bikevol$segment_na, "<br/>", 
   "Bike_Vol: ", round(sf_bikevol$vol_6ydaily_bike), sep="") %>%
  lapply(htmltools::HTML) 

pop_ped <- paste(
   "State: ", sf_pedvol$state, "<br/>", 
   "Segment: ", sf_pedvol$segment_na, "<br/>", 
   "Ped_Vol: ", round(sf_pedvol$vol_6ydaily_ped), sep="") %>%
  lapply(htmltools::HTML) 

tag.map.title <- tags$style(HTML(".leaflet-control.map-title 
{transform: translate(50%,50%);
position: fixed !important;
left: 45%;
text-align: center;
padding-left: 10px;
padding-right: 10px;
background: rgba(255,255,255,0.75);
font-weight: bold;
font-size: 36px;
}
"))

title_dcva <- tags$div(
  tag.map.title, HTML("Daily Bike & Ped Volume: 2018-2023")
)
title_or <- tags$div(
  tag.map.title, HTML("Daily Bike Volume: 2018-2023")
)

### MAP - DC & VA (zoom at DC Area)

# Creating Volume Map:DC&VA
m_volume_dcva <- leaflet() %>% 
   addTiles() %>% 
   setView( lng = -77.0369, lat = 38.9072, zoom = 12) %>% 
   addProviderTiles("Esri.WorldTopoMap")%>%
  
   # Add 2 marker groups: Volume
     addCircleMarkers(data=sf_pedvol, lng=~long , lat=~lat, 
                      radius=~vol_6ydaily_ped/100,
                      fillColor="purple", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Pedestrian Volume",
                      label = pop_ped,
                      labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), 
                                                   textsize = "13px", direction = "auto")) %>% 
       addCircleMarkers(data=sf_bikevol, lng=~long , lat=~lat, 
                      radius=~vol_6ydaily_bike/100, 
                      fillColor="blue", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Bike Volume",
                      label = pop_bike,
                      labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), 
                                                   textsize = "13px", direction = "auto")) %>%


   # Add the control widget
     addLayersControl(overlayGroups = c("Bike Volume","Pedestrian Volume"),  
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addControl(title_dcva, position = "bottomright")

m_volume_dcva 


# MAP - OR(zoom at Portland Metropolitan Area)
m_volume_or <- leaflet() %>% 
   addTiles() %>% 
   setView( lng = -122.676483, lat = 45.523064, zoom = 12) %>% 
   addProviderTiles("Esri.WorldTopoMap") %>%
  
   # Add 1 marker: Volume
     addCircleMarkers(data=sf_bikevol, lng=~long , lat=~lat, 
                      radius=~vol_6ydaily_bike/30, 
                      fillColor="blue", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Bike Volume",
                      label = pop_bike,
                      labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), 
                                                   textsize = "13px", direction = "auto"))%>% 

   # Add the control widget
     addLayersControl(overlayGroups = c("Bike Volume"),  
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addControl(title_or, position = "bottomright")

m_volume_or
```

# 3. Where do Bike/Ped Crash Occur?
## 3.1. Population Density & Bike/Ped Crash
```{r echo=FALSE, warning=FALSE}

### MAP Crash & Pop Density - DC & VA (zoom at DC Area)

# Clean the data object
sf_popdenhalf_dcva <- sf_popdenhalf_dcva %>%
  mutate(popden_sqmi = ifelse(popden_sqm == 0, NA, round(popden_sqm/ 1000, 2)))

library(RColorBrewer)
mybins_dcva_popden <- c(0, 5, 10, 20, 30, 50, 100, Inf)
mypalette_dcva <- colorBin(
  palette = "YlGn", domain = sf_popdenhalf_dcva$popden_sqmi,
  na.color = "transparent", bins = mybins_dcva_popden
)

# Creating Crash & Pop Density Map:DC&VA
m_crash_dcva <- leaflet(sf_popdenhalf_dcva) %>% 
   addTiles() %>% 
   setView( lng = -77.0369, lat = 38.9072, zoom = 11) %>% 
   addProviderTiles("Esri.WorldTopoMap")%>%
  
   # Add choropleth groups: Population Density
  addPolygons(stroke = FALSE, fillOpacity = 0.8,
    smoothFactor = 0.2, color = ~mypalette_dcva(popden_sqmi)) %>%
    
  # Add the palette legend
   leaflet::addLegend(
     pal = mypalette_dcva, values = ~popden_sqmi, 
     title = "Pop. Density (K/sqmi)", position = "bottomleft", 
     opacity = 0.8)  %>%

  # Add 2 marker groups: Crash
     addCircleMarkers(data=sf_bikecrash_dcva, 
                      radius=1.5, 
                      fillColor="red", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Bike Crash") %>%
     addCircleMarkers(data=sf_pedcrash_dcva, 
                      radius=1.5,
                      fillColor="orange", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Pedestrian Crash") %>% 

   # Add the control widget
     addLayersControl(overlayGroups = c("Bike Crash", "Pedestrian Crash"),  
                   options = layersControlOptions(collapsed = FALSE))

m_crash_dcva 

### MAP Crash & Pop Density - OR (zoom at DC Area)

# Clean the data object
sf_popdenhalf_or <- sf_popdenhalf_or %>%
  mutate(popden_sqmi = round(popden/1000, 2))

mybins_or_popden <- c(0, 5, 10, 20, 30, 50, 100, Inf)
mypalette_or <- colorBin(
  palette = "YlGn", domain = sf_popdenhalf_or$popden_sqmi,
  na.color = "transparent", bins = mybins_or_popden
)

# MAP - OR(zoom at Portland Metropolitan Area)
m_crash_or <- leaflet(sf_popdenhalf_or) %>% 
   addTiles() %>% 
   setView( lng = -122.676483, lat = 45.523064, zoom = 11) %>% 
   addProviderTiles("Esri.WorldTopoMap") %>%
  
  # Add choropleth groups: Population Density
  addPolygons(stroke = FALSE, fillOpacity = 0.8,
    smoothFactor = 0.2, color = ~mypalette_or(popden_sqmi)) %>%
    
  # Add the palette legend
   leaflet::addLegend(
     pal = mypalette_or, values = ~popden_sqmi, 
     title = "Pop. Density (K/sqmi)", position = "bottomleft", 
     opacity = 0.8)  %>%
  
  # Add 2 marker : Crash
     addCircleMarkers(data=sf_bikecrash_or, 
                      radius=1.5, 
                      fillColor="red", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Bike Crash") %>%

    
   # Add the control widget
     addLayersControl(overlayGroups = c("Bike Crash"),  
                   options = layersControlOptions(collapsed = FALSE))
m_crash_or
```
## 3.2. Bike/Ped Volume & Bike/Ped Crash

```{r echo=FALSE, warning=FALSE}

### MAP Crash & VOLUME - DC & VA (zoom at DC Area)

# Clean the data object

sf_bikepedvol_vol_dcva <- sf_bikepedvol_vol_dcva %>%
  mutate(bike_vol_fin = ifelse(bike_vol_f == 0, NA, round(bike_vol_f, 2))) %>%
  mutate(ped_vol_fin = ifelse(ped_vol_fi == 0, NA, round(ped_vol_fi, 2)))

library(RColorBrewer)
mybins_dcva_bikevol <- c(0, 50, 100, 150, 200, 250, 300, Inf)
mypalette_dcva_bike <- colorBin(
  palette = "Blues", domain = sf_bikepedvol_vol_dcva$bike_vol_fin,
  na.color = "transparent", bins = mybins_dcva_bikevol
)

mybins_dcva_pedvol <- c(0, 100, 200, 300, 400, 500, 600, Inf)
mypalette_dcva_ped <- colorBin(
  palette = "BuPu", domain = sf_bikepedvol_vol_dcva$ped_vol_fin,
  na.color = "transparent", bins = mybins_dcva_pedvol
)


# Creating Crash & Pop Density Map:DC&VA
m_crash_vol_dcva <- leaflet(sf_bikepedvol_vol_dcva) %>% 
   addTiles() %>% 
   setView( lng = -77.0369, lat = 38.9072, zoom = 11) %>% 
   addProviderTiles("Esri.WorldTopoMap")%>%
  
   # Add choropleth groups: Population Density
  addPolygons(stroke = FALSE, fillOpacity = 0.8,
    smoothFactor = 0.2, color = ~mypalette_dcva_bike(bike_vol_fin),group="Bike Volume") %>%
  
    addPolygons(stroke = FALSE, fillOpacity = 0.8,
    smoothFactor = 0.2, color = ~mypalette_dcva_ped(ped_vol_fin),group="Pedestrian Volume") %>%
    
  # Add the palette legend
   leaflet::addLegend(
     pal = mypalette_dcva_bike, values = ~bike_vol_fin, 
     title = "Bike Volume", position = "bottomleft", 
     opacity = 0.8,group="Bike Volume")  %>%
  
     leaflet::addLegend(
     pal = mypalette_dcva_ped, values = ~ped_vol_fin, 
     title = "Pedestrian Volume", position = "bottomleft", 
     opacity = 0.8,group="Pedestrian Volume")  %>%

  # Add 2 marker groups: Crash
     addCircleMarkers(data=sf_bikecrash_dcva_volume, 
                      radius=1.5, 
                      fillColor="red", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Bike Crash") %>%
     addCircleMarkers(data=sf_pedcrash_dcva_volume, 
                      radius=1.5,
                      fillColor="orange", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Pedestrian Crash") %>% 

   # Add the control widget
     addLayersControl(overlayGroups = c("Bike Crash", "Pedestrian Crash","Bike Volume","Pedestrian Volume"),  
                   options = layersControlOptions(collapsed = FALSE))

m_crash_vol_dcva 

### MAP Crash & Volume - OR (zoom at DC Area)

# Clean the data object
sf_bikevol_half_or <- sf_bikevol_half_or %>%
  mutate(bike_vol_fin = round(bike_vol_f, 2))

mybins_or_bike <- c(0, 50, 100, 200, 300, 400, 500, Inf)
mypalette_or_bike <- colorBin(
  palette = "Blues", domain = sf_bikevol_half_or$bike_vol_fin,
  na.color = "transparent", bins = mybins_or_bike
)

# MAP - OR(zoom at Portland Metropolitan Area)
m_crash_or_bike <- leaflet(sf_bikevol_half_or) %>% 
   addTiles() %>% 
   setView( lng = -122.676483, lat = 45.523064, zoom = 12) %>% 
   addProviderTiles("Esri.WorldTopoMap") %>%
  
  # Add choropleth groups: Population Density
  addPolygons(stroke = FALSE, fillOpacity = 0.8,
    smoothFactor = 0.2, color = ~mypalette_or_bike(bike_vol_fin),group="Bike Volume") %>%
    
  # Add the palette legend
   leaflet::addLegend(
     pal = mypalette_or_bike, values = ~bike_vol_fin, 
     title = "Bike Volume", position = "bottomleft", 
     opacity = 0.8,group="Bike Volume")  %>%
  
  # Add 2 marker : Crash
     addCircleMarkers(data=sf_bikecrash_or_volume, 
                      radius=1.5, 
                      fillColor="red", 
                      stroke = FALSE, 
                      fillOpacity = 0.8, 
                      group="Bike Crash") %>%

    
   # Add the control widget
     addLayersControl(overlayGroups = c("Bike Crash","Bike Volume"),  
                   options = layersControlOptions(collapsed = FALSE))
m_crash_or_bike
```

### 3.2.1 Bike Volume & Bike Crash by Land Use Type

```{r echo=FALSE, warning=FALSE}
ggplot(data_segment_ave) +
  aes(x=land_use, y=vol_6ydaily_bike, color = land_use) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  ggtitle("Bike Volume by Land Use Type") +
  theme_light()+
  xlab("Land Use") + 
  ylab("Bike Volume")
```

```{r echo=FALSE, warning=FALSE}
ggplot(data_segment_ave) +
  aes(x=land_use, y=bikecr_yrave, color = land_use) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  ggtitle("Bike-involved Crash by Land Use Type") +
  theme_light()+
  xlab("Land Use") + 
  ylab("Bike Crash")
```

### 3.2.2. Pedestrian Volume & Pedestrian Crash by Land Use Type
```{r echo=FALSE, warning=FALSE}
ggplot(data_segment_ave) +
  aes(x=land_use, y=vol_6ydaily_ped, color = land_use) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  ggtitle("Pedestrian Volume by Land Use Type") +
  theme_light()+
  xlab("Land Use") + 
  ylab("Pedestrian Volume")

```

```{r echo=FALSE, warning=FALSE}
ggplot(data_segment_ave) +
  aes(x=land_use, y=pedcr_yrave, color = land_use) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  ggtitle("Pedestrian-involved Crash by Land Use Type") +
  theme_light()+
  xlab("Land Use") + 
  ylab("Pedestrian-involved Crash")
```


# 4. When do Bike/Ped Crash Occur?
## 4.1. Yearly Trend
### 4.1.1. Bike Volume & Bike Crash
```{r echo=FALSE, warning=FALSE}
ggplot(data_state_year) +
  aes(x=year, y=(bike_vol), color=state) +
    geom_line(size=1) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Bike Volume by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Year") + ylab("Bike Volume")+
      scale_y_continuous(limits = c(0,500))
```

```{r echo=FALSE, warning=FALSE}
ggplot(data_state_year) +
  aes(x=year, y=num_bikecrash_persegmentid, color=state) +
    geom_line(size=1) +
    geom_point(shape=21, color="black", fill="red", size=2.5) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Bike-involved Crashes by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Year") + ylab("Bike Crashes (n/number of Permanent Count Location)")+
   ylim(0,12)
```

### 4.1.2. Pedestrian Volume & Pedestrian Crash
```{r echo=FALSE, warning=FALSE}
# Custom Y-axis labels
data_ped_year <- data_state_year %>% filter(state != "OR")
ggplot(data_ped_year) +
  aes(x=year, y=(ped_vol), color=state) +
    geom_line(size=1) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Pedestrian Volume by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Year") + ylab("Average Pedestrian Volume")+
    scale_y_continuous(limits= c(0,1500))
```

```{r echo=FALSE, warning=FALSE}
ggplot(data_ped_year) +
  aes(x=year, y=num_pedcrash_persegmentid, color=state) +
    geom_line(size=1) +
    geom_point(shape=21, color="black", fill="red", size=2.5) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Pedestrian-involved Crashes by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Year") + ylab("Pedestrian Crashes (n/number of Permanent Count Location)")+
   ylim(0,17)
```

## 4.2. Monthly Trend

### 4.2.1. Bike Volume & Bike Crash
```{r echo=FALSE, warning=FALSE}
ggplot(data_state_month) +
  aes(x=month, y=(bike_vol_ave_bysegment), color=state) +
    geom_line(size=1) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Bike Volume by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Month") + ylab("Bike Volume")+
      scale_y_continuous(limits = c(0,600))+
   scale_x_continuous(
    breaks = seq_along(month.name), 
    labels = month.name
  ) +theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

```{r echo=FALSE, warning=FALSE}
ggplot(data_state_month) +
  aes(x=month, y=bike_crash_ave_bysegment, color=state) +
    geom_line(size=1) +
    geom_point(shape=21, color="black", fill="red", size=2.5) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Bike-involved Crashes by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Month") + ylab("Bike Crashes (n/number of Permanent Count Location)")+
   scale_y_continuous(limits = c(0,36))+   
   scale_x_continuous(
    breaks = seq_along(month.name), 
    labels = month.name
  ) +theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

### 4.2.2. Pedestrian Volume & Pedestrian Crash
```{r echo=FALSE, warning=FALSE}
data_ped_month <- data_state_month %>% filter(state != "OR")

ggplot(data_ped_month) +
  aes(x=month, y=(ped_vol_ave_bysegment), color=state) +
    geom_line(size=1) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Pedestrian Volume by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Month") + ylab("Ped Volume")+
      scale_y_continuous(limits = c(0,1400))+
   scale_x_continuous(
    breaks = seq_along(month.name), 
    labels = month.name
  ) +theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

```{r echo=FALSE, warning=FALSE}
ggplot(data_ped_month) +
  aes(x=month, y=ped_crash_ave_bysegment, color=state) +
    geom_line(size=1) +
    geom_point(shape=21, color="black", fill="red", size=2.5) +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Pedestrian-involved Crashes by Permanent Count Location: 2018-2023") +
      theme_light()+
       xlab("Month") + ylab("Ped Crashes (n/number of Permanent Count Location)")+
   scale_y_continuous(limits = c(0,40))+   
   scale_x_continuous(
    breaks = seq_along(month.name), 
    labels = month.name
  ) +theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

# 5. Speed

## 5.1. Characteristics of Speed by State
```{r echo=FALSE, include=FALSE}
speed_year <- group_by(speed, segment_area_id, state, year)
speed_year <- summarise(speed_year,
                        speed_50th=mean(segment_50thmed),
                        speed_85th=mean(segment_85thmed),
                        speed_max=max(segment_max))
```

```{r echo=FALSE, warning=FALSE}
ggplot(speed_year) +
  aes(x=state, y=speed_85th, color = state) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.7, option="A") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Speed (85th) by State") +
  theme_light()+
  xlab("State") + 
  ylab("Speed 85th (mph)")

```

## 5.2. Relationship between Volume,Speed, and Crash
```{r echo=FALSE, warning=FALSE}
data_6_2 <- bikevol_crash_speed_segment_year %>% filter (bikevol<12000)
ggplot(data_6_2, aes(y=bikevol, x=speed_50th, shape=state, color=state)) + 
    geom_point(size=3) +
    ggtitle("Bike Volume & Vehicle Speed (50th)") +
    ylab("Bike Volume") + xlab("Vehicle Speed 50th (mph)")+
    theme_ipsum()

ggplot(data_6_2, aes(x=speed_85th, y=bikecrash, shape=state, color=state)) + 
    geom_point(size=3) +
    ggtitle("Bike Crash & Vehicle Speed (85th)") +
    xlab("Vehicle Speed 85th (mph)") + ylab("Bike Crashes")+
    theme_ipsum()


bike3d_speed_85 <- plot_ly(data_6_2, x = ~bikevol, y = ~speed_85th, z = ~bikecrash,
               marker = list(color = ~bikecrash, colorscale = c('mistyrose', 'red4'), showscale = TRUE)) %>% add_markers()
bike3d_speed_85 <- bike3d_speed_85 %>% layout(scene = list(xaxis = list(title = 'Bike Volume'),
                                   yaxis = list(title = 'Vehicle Speed'),
                                   zaxis = list(title = 'Bike Crash')),
                      annotations = list(
                        x = 1.13,
                        y = 1.05,
                        text = 'Bike Crash number within half-mile buffer area from the Permanent Bike/Ped Count Location',
                        xref = 'paper',
                        yref = 'paper',
                        showarrow = FALSE
                        ))
bike3d_speed_85
```

```{r echo=FALSE, warning=FALSE}

data_6_2 <- pedvol_crash_speed_segment_year %>% filter (pedvol<5200)

ggplot(data_6_2, aes(y=pedvol, x=speed_50th, shape=state, color=state)) + 
    geom_point(size=3) +
    ggtitle("Pedestrian Volume & Vehicle Speed (50th)") +
    ylab("Pedestrian Volume") + xlab("Vehicle Speed 50th (mph)")+
    theme_ipsum()


ggplot(data_6_2, aes(x=speed_85th, y=pedcrash, shape=state, color=state)) + 
    geom_point(size=3) +
    ggtitle("Pedestrian Crash & Vehicle Speed (85th)") +
    xlab("Vehicle Speed 85th (mph)") + ylab("Pedestrian Crashes")+
    theme_ipsum()


ped3d_speed_85 <- plot_ly(data_6_2, x = ~pedvol, y = ~speed_85th, z = ~pedcrash,
               marker = list(color = ~pedcrash, colorscale = c('mistyrose', 'red4'), showscale = TRUE)) %>% add_markers()
ped3d_speed_85 <- ped3d_speed_85 %>% layout(scene = list(xaxis = list(title = 'Pedestrian Volume'),
                                   yaxis = list(title = 'Vehicle Speed'),
                                   zaxis = list(title = 'Pedestrian Crash')),
                      annotations = list(
                        x = 1.13,
                        y = 1.05,
                        text = 'Pedestrian Crash number within half-mile buffer area from the Permanent Bike/Ped Count Location',
                        xref = 'paper',
                        yref = 'paper',
                        showarrow = FALSE
                        ))
ped3d_speed_85
```